Paper_Cartama_Microbial_ecology

### 2023-04-01 R/Paper_plots_microbial_ecology.Rmd
before last revision

```{r Packages}

library(tidyverse)
library(gridExtra)
library(tibble)
library(ggplot2)
library(extrafont)
loadfonts(device = "win")
library(phyloseq)
library(scales)   
library(ggpubr)
library(readxl)
library(vegan)
library(biomformat)
library(patchwork)

path<-"C:/Users/Dru/Desktop/Cartama_paper/files/MicrobialEcology"

@@

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
```

@@

```{r alfa df}
Alfa <- read_excel(paste(path,"/Alfa.xlsx",sep=""))
head(Alfa)
```
Ploting the alpha_diversity metrics

@@

c <-  ggplot(Alfa, aes(x = Damage_level, y = Chao)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
c <- c + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
c <- c +  xlab("") + ylab("Chao")+ylim(c(0, max(Alfa$Chao)*1.1))
c <- c + theme(axis.title =element_text(size=14,family = 'serif'))
c <- c +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
c <- c + theme(strip.text.x = element_text(size = 14,family = 'serif'))
c <- c + geom_signif(test="t.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

s <-  ggplot(Alfa, aes(x = Damage_level, y = Shannon_entropy)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
s <- s + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
s <- s +  xlab("") + ylab("Shannon entropy")+ylim(c(0, max(Alfa$Shannon_entropy)*1.1))
s <- s + theme(axis.title =element_text(size=14,family = 'serif'))
s <- s +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
s <- s + theme(strip.text.x = element_text(size = 14,family = 'serif'))
s <- s + geom_signif(test="t.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

r <-  ggplot(Alfa, aes(x = Damage_level, y = Riqueza)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
r <- r + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
r <- r +  xlab("") + ylab("Richness")+ylim(c(0, max(Alfa$Riqueza)*1.1))
r <- r + theme(axis.title =element_text(size=14,family = 'serif'))
r <- r +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
r <- r + theme(strip.text.x = element_text(size = 14,family = 'serif'))
r <- r + geom_signif(test="t.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

f <-  ggplot(Alfa, aes(x = Damage_level, y = Faith_pd)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
f <- f + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
f <- f +  xlab("") + ylab("Faith pd")+ylim(c(0, max(Alfa$Faith_pd)*1.1))
f <- f + theme(axis.title =element_text(size=14,family = 'serif'))
f <- f +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
f <- f + theme(strip.text.x = element_text(size = 14,family = 'serif'))
f <- f + geom_signif(test="t.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

 alpha_plot<-grid.arrange(r + labs(tag = 'A'),s + labs(tag = 'B'), f + labs(tag = 'C'), ncol=3,nrow=1)
```

@@

```{r uploading the CCS.biom and tree.nwk}
biomfile <- import_biom(paste(path,"/css_eff.biom",sep=""))
biomfile_mapfile <- import_qiime_sample_data(paste(path,"/metadata_real.txt",sep=""))
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)

otu.table <- otu_table(biomfile_merged)
taxa <- tax_table(biomfile_merged)

tree <- read_tree(paste(path,"/tree.nwk",sep=""))

@@

>>>>

```{r Packages, message=FALSE, warning=FALSE}

library(tidyverse)
library(gridExtra)
library(tibble)
library(ggplot2)
library(extrafont)
loadfonts(device = "win")
library(phyloseq)
library(scales)   
library(ggpubr)
library(readxl)
library(vegan)
library(biomformat)
library(patchwork)
library(car)
library(ggimage)
library(ggtree)
library(TDbook)
library(aplot)
library(lemon)
library(grid, lib.loc = "C:/Program Files/R/R-4.2.0/library")
setwd("C:/Users/Dru/Desktop/Cartama_paper/microbial-ecology-avocado/Resultados_all")
```

@@

 plot<-grid.arrange(p + labs(tag = 'A'),p1 + labs(tag = 'B'), ncol=1,nrow=2)
ggsave("C:/Users/Dru/Desktop/Cartama_paper/SupplementaryFig1.jpg",plot=plot,dpi = 300,width = 7,height = 7,units = c("in"))
```
## Rarefraction curve plots

```{r}

#Curvas de rarefacci?n antes de normalizacion
biomfile_before <- import_biom("ASVs_table6_dada2R.biom")
biomfile_after <- import_biom("ASVs_table9_dada2R.biom")
head(otu_table(biomfile))
head(tax_table(biomfile))

otu.table_before <- otu_table(biomfile_before)
t.OTU.table_before <- t(otu.table_before)

otu.table_after <- otu_table(biomfile_after)
t.OTU.table_after <- t(otu.table_after)
#head(otu_table(biomfile_before))
#head(tax_table(biomfile))

col <- c("black", "darkred", "forestgreen", "orange", "blue", "yellow", "hotpink")
lty <- c(rep("solid",6),rep( "dashed",6), rep( "longdash", 6), rep("dotdash",6))
#par(mar=c(5,4,2,2))
#par(mgp=c(2.3,0.5,0))
jpeg("rarre.jpg", units="in", width=7, height=6, res=300)
par(mfrow=c(1,2))
p<-rarecurve(t.OTU.table_before, col="black", step = 1000, cex = 0.8, lwd=1, label = FALSE,xlab = "Sample size", ylab = "Specie Number", ylim=c(0, 680),family="serif",cex.lab=1.4, cex.axis=1.2,lty=lty)

text(-14800, 760, "A", family = "serif", xpd=NA, cex = 1.2 )
p1<-rarecurve(t.OTU.table_after, col="black", step = 1000, cex = 0.8, lwd=1, label = FALSE,xlab = "Sample size", ylab = "Specie Number", ylim=c(0, 680),family="serif",cex.lab=1.4, cex.axis=1.2,lty=lty)

text(-6000, 760, "B", family = "serif", xpd=NA, cex = 1.2 )
dev.off()

#The rarefaction curves are evaluated using the interval of step sample sizes
#Graphical Parameters: https://www.statmethods.net/advgraphs/parameters.html

jpeg("rarre.jpg",, units="in", width=5, height=5, res=300)
dev.off()
```

@@

```{r alfa df}
path<-"C:/Users/Dru/Desktop/Cartama_paper/microbial-ecology-avocado/Resultados_all/Alfa/"
Alfa <- read_excel(paste(path,"Alfa.xlsx",sep=""))
head(Alfa)
```
```{r alfa analisis}
c <- subset(Alfa, Alfa$Farm == "Sinai")$Chao
s <- subset(Alfa, Alfa$Farm == "Sinai")$Shannon_entropy
r <- subset(Alfa, Alfa$Farm == "Sinai")$Riqueza
f <- subset(Alfa, Alfa$Farm == "Sinai")$Faith_pd
e <- subset(Alfa, Alfa$Farm == "Sinai")$Evenness

damage <-subset(Alfa, Alfa$Farm == "Escondida")$Damage_level

#names(shannon)
#Check ANOVA assumptions
#Homogeneity of variance
leveneTest(c ~ damage)  #p-value > 0.05 
#Shapiro-Wilk test
shapiro.test(c) #p-value > 0.05 is normal
#Independence
durbinWatsonTest(lm(c ~ damage)) #p-value > 0.05 they are independent

t.test(r~damage)
kruskal.test(s~damage)
summary(glm(c~damage, family = quasipoisson()))
plot(glm(c~damage, family = poisson()))
plot(lm(c~damage))
```
Ploting the alpha_diversity metrics

@@

c <-  ggplot(Alfa, aes(x = Damage_level, y = Chao)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
c <- c + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
c <- c +  xlab("") + ylab("Chao")+ylim(c(0, max(Alfa$Chao)*1.1))
c <- c + theme(axis.title =element_text(size=14,family = 'serif'))
c <- c +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
c <- c + theme(strip.text.x = element_text(size = 14,family = 'serif'))
c <- c + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

s <-  ggplot(Alfa, aes(x = Damage_level, y = Shannon_entropy)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
s <- s + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
s <- s +  xlab("") + ylab("Shannon entropy")+ylim(c(0, max(Alfa$Shannon_entropy)*1.1))
s <- s + theme(axis.title =element_text(size=14,family = 'serif'))
s <- s +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
s <- s + theme(strip.text.x = element_text(size = 14,family = 'serif'))
s <- s + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

r <-  ggplot(Alfa, aes(x = Damage_level, y = Riqueza)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
r <- r + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
r <- r +  xlab("") + ylab("Richness")+ylim(c(0, max(Alfa$Riqueza)*1.1))
r <- r + theme(axis.title =element_text(size=14,family = 'serif'))
r <- r +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
r <- r + theme(strip.text.x = element_text(size = 14,family = 'serif'))
r <- r + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

f <-  ggplot(Alfa, aes(x = Damage_level, y = Faith_pd)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
f <- f + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
f <- f +  xlab("") + ylab("Faith pd")+ylim(c(0, max(Alfa$Faith_pd)*1.1))
f <- f + theme(axis.title =element_text(size=14,family = 'serif'))
f <- f +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
f <- f + theme(strip.text.x = element_text(size = 14,family = 'serif'))
f <- f + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

###evenness
e <-  ggplot(Alfa, aes(x = Damage_level, y = Evenness)) + 
      geom_bar(stat="summary", color="black", position=position_dodge(),
               fill=rep(c("gray90","gray70"),2),fun = mean) +
      stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), geom="errorbar",
                   color="black", width=0.2) 
e <- e + geom_jitter()+ facet_wrap(~Farm)+ theme_classic() 
e <- e +  xlab("") + ylab("Evenness")+ylim(c(0, max(Alfa$Evenness)*1.1))
e <- e + theme(axis.title =element_text(size=14,family = 'serif'))
e <- e +  theme(axis.text.x = element_text(face="bold",size=12, angle=0,
                                             family ='serif'),
                  axis.text.y = element_text(face="bold",size=12, angle=0,family =
                                               'serif'))
e <- e + theme(strip.text.x = element_text(size = 14,family = 'serif'))
e <- e + geom_signif(test="kruskal.test", comparisons = list(c("Mild", "Severe")), map_signif_level = T)

@@

 alpha_plot<-grid.arrange(s + labs(tag = 'A'),c + labs(tag = 'B'), e + labs(tag = 'C'),f + labs(tag = 'D'), ncol=2,nrow=2)
ggsave("Figure2_alfa_new.jpg",plot=alpha_plot,dpi = 300,width = 6,height = 6,units = c("in"))
```

@@

```{r uploading the CCS.biom and tree.nwk}

path<-"C:/Users/Dru/Desktop/Cartama_paper/microbial-ecology-avocado/Resultados_All"
biomfile <- import_biom(paste(path,"/css_eff.biom",sep=""))
biomfile_mapfile <- import_qiime_sample_data(paste(path,"/metadata_real.txt",sep=""))
tree <- read_tree(paste(path,"/tree.nwk",sep=""))

biomfile_mapfile_sinai<- subset(biomfile_mapfile,biomfile_mapfile$Farm1=="Sinai")
biomfile_mapfile_escondida<-subset(biomfile_mapfile,biomfile_mapfile$Farm1=="Escondida")

biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile_escondida)
colnames(tax_table(biomfile_merged)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)
otu.table <- otu_table(biomfile_merged)
taxa <- tax_table(biomfile_merged)

@@

```{r CAP}
agricultural.ord<- ordinate(merge2, "CAP", "wunifrac", ~Farm*Lenticel_disease_status)

CAP = agricultural_plot  +
  geom_point(aes(color =Farm,shape = Farm1), size = 3) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white", colour = "black",size = 0.5, linetype = 1),
        axis.ticks = element_line(colour = "black"),
        axis.title.x = element_text(face="plain",family = "serif", colour="black", size=12),
        axis.title.y = element_text(face="plain",family = "serif", colour="black", size=12),
        axis.text.x  = element_text(face="plain",family = "serif", colour = "black", angle=360, vjust=0.5, size=10),
        axis.text.y  = element_text(face="plain",family = "serif", colour = "black", angle=360, vjust=0.5, size=10),
        panel.background = element_rect(fill = "white", colour = "black", linetype = 1, size = 0.5),
        panel.border = element_blank(), legend.title = element_blank(), legend.text=element_text(size=0.1,family = "serif"), plot.margin = margin(0.755, 1, 1, 0.3, "cm"))  +
  scale_colour_manual(values=c("gray85","gray55","gray85","gray55"),labels =
                        c("Escondida_Mild","Escondida_Severe",
                          "Sinai_Mild","Sinai_Severe")) + 
  stat_ellipse(type = "t", size=0.5) + labs(x = "CAP (23.7%)", y= "CAP (45.8%)") + theme(legend.position = "right",legend.text = element_text(size=10))
CAP
ggsave("Figure4_new.jpg",plot=CAP,dpi = 300,width = 6,height = 5,units = c("in"))

*Fig 4* Fungal communities’ structure for fruits of avocado cv. Hass fruits with mild and severe lenticel damages collected from the La Escondida and El Sinai farms during the traviesa harvest of 2020. Cumulative sum scaling (CSS) transformed reads were used to calculate weighted unifrac distance. Shown is the bidimensional plane of the canonical analysis of principal coordinates (CAP) constrained by the farm and the strenght of the damage (69.5 % of the overall variance; p-value: 0.001 acording to the permanova test). Each point represents a fruit's fungal community, the shapes show the farm (La Escondida (circles) and El Sinai (triangles)), and the colors the damage strength (Severe (dark gray) and Mild (light gray)) of the fruit. The ellipses are the 95 % confidence level ellipses.

```{r}

#agricultural.ord<- ordinate(merge2, "CAP", "wunifrac", ~Farm1*Lenticel_damage_status)

#PERMANOVA
perm_anova.agricultural.ord1 <- anova.cca(ordinate(merge2, "CAP", "wunifrac", ~Lenticel_disease_status))
#perm_anova.agricultural2.ord <- permutest(agricultural.ord)

# P-value
print(as.data.frame(perm_anova.agricultural.ord)["Pr(>F)"])
```

po<-grid.arrange(PCoA+labs(tag = "A"),CAP+  geom_text(x=-0.65, y=0.69,label="B",size=4.5, col= "black") +coord_cartesian( clip = "off"),legend, ncol=2,nrow = 2, 
             layout_matrix = rbind(c(1,2), c(3,3)),
             widths = c(2.7, 2.7), heights = c(2.5, 0.2))
ggsave("C:/Users/Dru/Desktop/Cartama_paper/Figure4.jpg",plot=po,dpi = 300,width = 7,height = 6,units = c("in"))
```
## Taxonomy bat plots
Loading the ASVs table

```{r data barplot}
biomfile <- import_biom(paste(path,"/ASVs_table9_dada2R.biom",sep=""))
##biomfile <- import_biom("ASVs_table6_dada2R.biom")
#biomfile_mapfile <- import_qiime_sample_data(paste(path,"/metadata_real.txt",sep=""))
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sample_data(biomfile_merged)

plot_colors <- c( "palegoldenrod","gray90", "plum4", "lightsteelblue","darkseagreen1",
                   "#CBD588", "#5F7FC7","#DA5724", "#8A7C64","gray30", "#508578", "#CD9BCD", "#AD6F3B",
                  "darkseagreen4","deepskyblue1","gray50" , "yellow2", "slateblue1",
                  "#673770","Khaki4", "#652926", "yellow3", "cadetblue3",
                   "lightsteelblue3", "honeydew3", "gray80","thistle4","yellow4",
                   "#5E738F", "#8A7C64","gray30", "lightpink3")
```

## Phylum barplot
```{r}
subset = biomfile_merged
rel.abun = transform_sample_counts(subset, function(x) x / sum(x))
glom <- tax_glom(rel.abun, taxrank = 'Phylum')
data_glom<- psmelt(glom) 
data_glom$Phylum <- as.character(data_glom$Phylum) 
data_glom$Phylum[data_glom$Abundance < 0.03] <- "< 3% abundance"
data_glom$Phylum[data_glom$Phylum == "unidentified"] <- "Not classified"
data_glom$Phylum[data_glom$Phylum == "undefined"] <-"Not classified"
data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Necrotic"] <-"Severe"
data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Healthy"] <-"Mild"

plot_colors <- c( "palegoldenrod","gray90", "plum4", "lightsteelblue","darkseagreen1")
Phylum <- ggplot(data_glom, aes(x = Lenticel_disease_status, y = Abundance/6, fill = reorder(Phylum, Abundance)))+
  geom_bar(stat = "identity") + facet_wrap(~Farm1) + scale_fill_manual(values =plot_colors) +
  ylab("Relative abundance") + xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white",  colour = "black",  size = 0.5, linetype = 1),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.title.y = element_text(face="plain", colour="black", size=13,family="serif"),
        axis.text.x  = element_text(face="plain", colour = "black", angle=30, vjust=0.6, size=13,family="serif"),
        axis.text.y  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=13,family="serif"),
        panel.background = element_rect(fill = "white", colour = "white", linetype = 1, size = 0.5),
        panel.border = element_blank()) + scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme(legend.title = element_blank(),legend.text = element_text(family="serif",size=12))+
  theme(strip.text.x = element_text(size = 13 , family = 'serif'))

print(Phylum)

df = data.frame(Treatment = data_glom$Farm, Phylum = data_glom$Phylum, Abundance = round(data_glom$Abundance,3)*100)
df = df[df$Phylum!="< 3% abundancia",]
df<-df[ order(df$Treatment,df$Phylum),]
mean<-tapply(df$Abundance, list(df$Phylum,df$Treatment),mean)
sd<-tapply(df$Abundance, list(df$Phylum,df$Treatment),sd)
sink("mean_phylum.txt")
print(mean)
sink()
sink("sd_phylum.txt")
print(sd)
sink()
```
## Class barplot
```{r}
subset = biomfile_merged
rel.abun = transform_sample_counts(subset, function(x) x / sum(x))
glom <- tax_glom(rel.abun, taxrank = 'Class')
data_glom<- psmelt(glom) 
data_glom$Class <- as.character(data_glom$Class) 
data_glom$Class[data_glom$Abundance < 0.03] <- "< 3% abundance"
data_glom$Class[data_glom$Class == "unidentified"] <- "Not classified"
data_glom$Class[data_glom$Class == "undefined"] <-"Not classified"

data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Necrotic"] <-"Severe"
data_glom$Lenticel_disease_status[data_glom$Lenticel_disease_status == "Healthy"] <-"Mild"

plot_colors <- c( "palegoldenrod","gray40","thistle1", "lightsteelblue","darkseagreen1",
                   "#CBD588", "plum4", "#652926","Khaki4", "yellow3", "cadetblue3",
                   "lightsteelblue3", "honeydew3",

                  "gray80", "yellow2", "slateblue1", "#5F7FC7","#DA5724", "#8A7C64",
                  "gray30", "#508578", "#CD9BCD", "gray80","#AD6F3B",
                  "darkseagreen4","deepskyblue1","gray50" , "yellow2", "slateblue1",
                  "#673770","thistle4","yellow4",
                   "#5E738F", "#8A7C64","gray30", "lightpink3")

Class <- ggplot(data_glom, aes(x = Lenticel_disease_status, y = Abundance/6, fill = reorder(Class, Abundance)))+
  geom_bar(stat = "identity") + facet_wrap(~Farm1) + scale_fill_manual(values =plot_colors) +
  ylab("Relative abundance") + xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        rect = element_rect(fill = "white",  colour = "black",  size = 0.5, linetype = 1),
        axis.line = element_line(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.title.y = element_text(face="plain", colour="black", size=13,family="serif"),
        axis.text.x  = element_text(face="plain", colour = "black", angle=30, vjust=0.6, size=13,family="serif"),
        axis.text.y  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=13,family="serif"),
        panel.background = element_rect(fill = "white", colour = "white", linetype = 1, size = 0.5),
        panel.border = element_blank()) + scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme(legend.title = element_blank(),legend.text = element_text(family="serif",size=12))+
  theme(strip.text.x = element_text(size = 13 , family = 'serif'))

print(Class)

df = data.frame(Treatment = data_glom$Farm, Class = data_glom$Class, Abundance = round(data_glom$Abundance,3)*100)
df = df[df$Class!="< 3% abundancia",]
df<-df[ order(df$Treatment,df$Class),]
mean<-tapply(df$Abundance, list(df$Class,df$Treatment),mean)
sd<-tapply(df$Abundance, list(df$Class,df$Treatment),sd)
sink("mean_Class.txt")
print(mean)
sink()
sink("sd_Class.txt")
print(sd)
sink()
```
```{r}

Phylum <- Phylum + theme(legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(0.2, 'cm'))

Class <- Class + theme(legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(0.2, 'cm'))

po<-ggarrange(ggarrange(Phylum+labs(tag= "A"),Class+labs(tag= "B"), nrow= 2)) 
po
ggsave("Fig4.jpg",plot=po,dpi = 300,width = 6,height = 8,units = c("in"))

### diferential abundance analysis

loading the data for diferencial abundance annalysis-Escondida

```{r escondida diferential abundance, message=FALSE, warning=FALSE}

biomfile <- import_biom("ASVs_enrich_Escondida.biom")
biomfile_mapfile <- import_qiime_sample_data("metadata_real1.txt")
tree <- read_tree("tree.nwk")
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c( "Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)
otu.table <- otu_table(biomfile_merged)

otu.table<-otu.table[!(otu.table[,1]<0.005&otu.table[,2]<0.005),]
#over 0.5%
taxa <- tax_table(biomfile_merged)
otutax = phyloseq(otu.table, taxa)
GP.chl = merge_phyloseq(otutax, sam, tree)

```{r making a df}

labels_new<-c("ASV22_Not_classified", "ASV57_g_Erythrobasidium", "ASV31_c_Cystobasidiomycetes", 
              "ASV30_g_Cystobasidium", "ASV53_g_Golubevia", "ASV25_g_Meira", "ASV75_g_Vishniacozyma",
              "ASV11_g_Vishniacozyma", "ASV15_g_Vishniacozyma", "ASV24_g_Setophoma","ASV34_g_Bacidina",
              "ASV18_g_Bacidina", "ASV6_g_Hyphozyma",  "ASV4_g_Cladosporium",  "ASV1_Not_classified",
              "ASV2_Not_classified",  "ASV12_Not_classified", "ASV3_Not_classified", "ASV7_Not_classified",
              "ASV42_Not_classified", "ASV41_Not_classified", "ASV60_g_Neodevriesia", "ASV9_Not_classified",
              "ASV10_o_Chaetothyriales", "ASV13_Not_classified")

tree_new<-phy_tree(GP.chl)
tree_new$tip.label<-labels_new
otu_table_new<-otu_table(GP.chl)
row.names(otu_table_new)<-labels_new
taxa_new<-tax_table(GP.chl)
row.names(taxa_new)<-labels_new
sample_data_new<-sample_data(GP.chl)
otutax_new = phyloseq(otu_table_new, taxa_new)
merge_new = merge_phyloseq(otutax_new, sample_data_new, tree_new)

df<- as.data.frame(otu_table_new)
df$Ratio <- df$Escondida_Severo/df$Escondida_Leve
df$ASVs<- phy_tree(tree_new)$tip.label
df$Enriched_in[df$Ratio < 1]<- "Mildly damaged fruits"
df$Enriched_in[df$Ratio >= 1]<- "Severely damaged fruits"

df1 <- data.frame(
  ASVs = phy_tree(tree_new)$tip.label, Ratio = df$Ratio, Enriched_in = df$Enriched_in 
)

df_anotations <- data.frame(node = c(1,28,35), 
                            taxa = c("Not lassified", "Basidiomycetes", "Ascomycetes"))

df_anotations_class <- data.frame(node = c(41,23,49,38,13,10,30,32,33,1), 
                                  Class = c("Dothideomycetes", "Not classified", "Eurotiomycetes", "Lecanoromycetes", 
                                            "Leotiomycetes", "Dothideomycetes", "Cystobasidiomycetes","Exobasidiomycetes",
                                            "Tremellomycetes", "Not classified"))

df_anotations_class$Class <- factor(df_anotations_class$Class, 
                                    levels = c("Dothideomycetes", "Not classified", "Eurotiomycetes", 
                                               "Lecanoromycetes", "Leotiomycetes", "Cystobasidiomycetes",
                                               "Exobasidiomycetes","Tremellomycetes"))
#p + geom_text(aes(label=node), hjust=-.3)

```
## Escondida plot

*Fig 6.* Taxonomy and relative abundance of the amplicon sequence variants (ASVs) enriched (p-value < 0.05) in the fungal communities of avocado cv. Hass fruits with different severities of lenticel damage (Mild and Severe) collected from the La Escondida during the traviesa harvest of 2020. Shown are the taxonomic relation (A) and the relative abundances (B) of the ASVs group by color according to the clade (Supplementary Table 5).

```{r ploting the data}

p <- ggtree(tree_new) + geom_tiplab(align=T, size=3) + xlim(-.1, 3)+ hexpand(.01)+ labs(tag = "A")

p <- p + geom_hilight(data=df_anotations_class, aes(node=node, fill=Class),  alpha=0.2,  extend=3, type = "roundrect")
p<-p+geom_nodepoint(aes(subset = node == 35, x = x - .07),size = 3, colour = 'brown3', shape = 19)
p<-p+geom_nodepoint(aes(subset = node == 28, x = x - 0.12),size = 3, colour = 'deepskyblue3', shape = 19)
p<- p +   theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10))

p1 <- ggplot(df1, aes(x = ASVs, y = log10(Ratio))) + geom_col(aes(fill=Enriched_in)) + theme_classic() + coord_flip() + labs(tag = "B")
p1 <- p1 + scale_fill_manual("Enriched in:",values= alpha(c("Mildly damaged fruits" = "#00A087B2", "Severely damaged fruits" ="#7E6148B2"),0.4))
p1 <- p1 + ylab("log10(Abundance severe/Abundance mild)") + xlab("") + guides(y = "none") + theme(axis.text.y=element_blank(),axis.ticks.y=element_blank())
p1<- p1+  theme(axis.title.x = element_text(face="plain", colour="black", size=11,family="serif"),
        axis.text.x  = element_text(face="plain",colour = "black", angle=360, vjust=0.5, size=10,family="serif")) +
  theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10))

po <- p1 %>% insert_left(p)

#####

data <- data.frame(Xdata = rnorm(2), Ydata = rnorm(2),Phylum =c("Ascomycete","Basidiomycete"))

gplot <- ggplot(data, aes(Xdata, Ydata, color = Phylum)) + geom_point(size = 3) +scale_color_manual(values = c("Ascomycete" = "brown3", "Basidiomycete" = "deepskyblue3"))
gplot <- gplot +   theme(legend.title = element_text(family="serif",size=12, face = "bold"),legend.text = element_text(family="serif",size=10),legend.key = element_rect(colour = NA, fill = NA))

jpeg("Fig5.jpg",width = 8,height = 6,units = c("in"), res=300) 
po
leg=g_legend(gplot)
wc_table_vp <- viewport(x = 0.74, y = 0.20, just = c("left", "top"),height = 0.1, width = 0.2)
pushViewport(wc_table_vp)
grid.draw(leg)

dev.off()

loading the data for diferencial abundance annalysis-Sinai

```{r Sinai diferential abundance, message=FALSE, warning=FALSE, paged.print=FALSE}

# biomfile <- import_biom("ASVs_enrich_Sinai.biom")

#biomfile_mapfile <- import_qiime_sample_data("metadata_real1.txt")
#tree <- read_tree("tree.nwk")
biomfile_merged <- merge_phyloseq(biomfile,biomfile_mapfile)
colnames(tax_table(biomfile_merged)) <- c( "Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
sam <-sample_data(biomfile_merged)
otu.table <- otu_table(biomfile_merged)

otu.table<-otu.table[!(otu.table[,1]<0.005&otu.table[,2]<0.005),]
#over 0.5%
taxa <- tax_table(biomfile_merged)
otutax = phyloseq(otu.table, taxa)
GP.chl = merge_phyloseq(otutax, sam, tree)

```{r making a df sinai}

labels_new<-c("ASV37_Not_classified", "ASV96_Not_classified","ASV33_g_Symmetrospora","ASV20_g_Symmetrospora","ASV31_c_Cystobasidiomycetes",
              "ASV54_c_Cystobasidiomycetes","ASV11_g_Vishniacozyma","ASV65_g_Colletotrichum", "ASV40_g_Geastrumia", "ASV14_g_Pseudocercospora", 
              "ASV4_g_Cladosporium",  "ASV2_Not_classified",  "ASV3_Not_classified",  "ASV7_Not_classified",  "ASV70_Not_classified",
              "ASV10_o_Chaetothyriales", "ASV13Not_classified", "ASV26_g_Aureobasidium")

tree_new<-phy_tree(GP.chl)
tree_new$tip.label<-labels_new
otu_table_new<-otu_table(GP.chl)
row.names(otu_table_new)<-labels_new
taxa_new<-tax_table(GP.chl)
row.names(taxa_new)<-labels_new
sample_data_new<-sample_data(GP.chl)
otutax_new = phyloseq(otu_table_new, taxa_new)
merge_new = merge_phyloseq(otutax_new, sample_data_new, tree_new)
